---
title: "Biodiversity analysis"
author: "Juliano Palacios-Abrantes"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, results='hide', message=FALSE, echo = F}

library(MyFunctions)

packages <- c(
      "readxl", # Read dataframe
      "tidyverse", # for all data wrangling and ggplot
      "janitor",
      "iNEXT",
      "googledrive",
      "googlesheets4"
      
)

my_lib(packages)
rm(packages)
# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)
```

# Load Data

```{r}
# spp_biodiv_df <- my_path("D", name = "spp_biodiversity_data.csv", read = T)


sheet_url <- "https://docs.google.com/spreadsheets/d/1ZnycZe_74bB-73LVCrBe56qEmBAa076bv4AFojcArzE/edit?gid=1581131140#gid=1581131140"

spp_biodiv_df <- read_sheet(sheet_url,
                            sheet = "Species_Occurence_Matrix",
                            col_types = "c"
) %>%
      janitor::clean_names() %>% 
      dplyr::mutate(across(cmnfi:species_occurence, ~ as.numeric(.x)))

fam_biodiv_df <- my_path("D", name = "fam_biodiversity_data.csv", read = T)

```

# Speceis curve Benthic pelagic 

```{r}

# Transform analysis into function to run by benthic/pelagic group
run_biodiv_analysis <- function(df_subset) {
      
      community_matrix_pres <- df_subset %>%
            pivot_longer(cols = cmnfi:united_states_geological_survey, # Note these must include all columns on data
                         names_to = "site", 
                         values_to = "value") %>%
            group_by(site, scientific_name) %>%
            summarise(presence = ifelse(sum(value) > 0, 1, 0), .groups = "drop") %>%
            reshape::cast(site ~ scientific_name, value = "presence")
      
      
      community_matrix_pres[is.na(community_matrix_pres)] <- 0
      
      inc_freq <- as.incfreq(as.matrix(t(community_matrix_pres)))
      iNEXT(inc_freq, q = 0, datatype = "incidence_freq", nboot = 1, endpoint = 22) # As pointed out by reviewer 
}


# Split and run analysis by benthic/pelagic group
results_list <- spp_biodiv_df %>%
      group_by(benthic_pelagic) %>%
      group_split() %>%
      set_names(unique(spp_biodiv_df$benthic_pelagic)) %>%
      map(run_biodiv_analysis)

# Extract size-based estimates and tag with group
benthic_df <- as.data.frame(results_list$Benthic$iNextEst$size_based) %>%
      mutate(Group = "Benthic")



pelagic_df <- as.data.frame(results_list$Pelagic$iNextEst$size_based) %>%
      mutate(Group = "Pelagic")

# Combine both into one data frame
Hills_q0_panel_df <- bind_rows(benthic_df, pelagic_df)


bio_spp <-
ggplot(Hills_q0_panel_df %>% filter(Method != "Observed")) +
      geom_line(aes(x = t,
                    y = qD,
                    color = Method,
                    lty = Group),
                # col = "coral2",
                cex = 1) +
      geom_point(data = Hills_q0_panel_df %>% filter(Method == "Observed"),
                 aes(x = t,
                     y = qD,
                     color = Method),
                 # col = "coral2",
                 cex = 2) +
      facet_wrap(~ Group, scales = "free", ncol = 2) +
      scale_color_viridis_d() +
      # xlab("") +
      xlab("Number of Sampling Units") +
      ylab("Species Diversity") +
      my_ggtheme_p(leg_pos = "right") +
      scale_y_continuous(breaks = seq(0, 500, 50)) +
      scale_x_continuous(breaks = seq(0, 22, 2)) +
      scale_linetype_manual(values = c("twodash", "solid"))


ggsave(bio_spp,
       filename = "benthic_pelagic_diversity.png",
       width = 9,
       height = 7)

```

## S(Observ)

In the context of biodiversity studies and using the iNEXT package, S(obs) refers to the observed species richness. This is the actual number of species observed in the samples collected. It is an important measure because it represents the baseline diversity observed in your data before any rarefaction or extrapolation.

```{r}

# For Benthic species
sb_obs <- results_list$Benthic$AsyEst$Observed[1] # 313
sb_se <- results_list$Benthic$AsyEst$Est_s.e.[1] #77.79394

# For Pelagic species
sp_obs <- results_list$Pelagic$AsyEst$Observed[1] # 137
sb_se <- results_list$Pelagic$AsyEst$Est_s.e.[1] #184.3977


bind_rows(
      # Benthic data
      results_list$Benthic$AsyEst %>% 
            mutate(category = c("Species Richness", "Shannon div.", "Simpson div"),
                   ,level = "secies",n = 11,
                   Benthic_Plagic = "Benthic") %>%
            select(category,everything()
            ),
      
      # Pelagic data
      results_list$Pelagic$AsyEst %>% 
            mutate(category = c("Species Richness", "Shannon div.", "Simpson div"),
                   ,level = "secies",n = 11,
                   Benthic_Plagic = "Pelagic") %>%
            select(category,everything()) 
      
) %>% 
      # View()
# Save dataset
# write_csv(
#       "spp_diversity_indices_bp.csv")

```

# Family curve

```{r}

community_matrix_pres <- fam_biodiv_df %>% 
      gather("site","value",cmnfi:united_states_geological_survey) %>%
      dplyr::group_by(site, family) %>%
      dplyr::summarise(presence = ifelse(sum(value) > 0, 1, 0)) %>%
      reshape::cast(.,  site ~ family, value = "presence")

community_matrix_pres[is.na(community_matrix_pres)] <-  0

inc_freq <- as.incfreq(as.matrix(t(community_matrix_pres)))
Hills_q_0_inc_CCZ <- iNEXT(inc_freq, q = 0, datatype = "incidence_freq", nboot = 1, endpoint = 22) # q and qdp is what plots
Hills_q0_inc_df <- as.data.frame(Hills_q_0_inc_CCZ$iNextEst$size_based)


bio_fam <-
ggplot(Hills_q0_inc_df %>% filter(Method != 'Observed')) +
      # geom_ribbon(aes(x = t, ymin = qD.LCL,ymax = qD.UCL, color = Method),
      #             alpha = 0.2, show.legend = FALSE) +      
      # for lines
      # geom_line(aes(x = t, y = qD, lty = Method), col = "coral2", cex = 1) +
      # geom_point(data = Hills_q0_inc_df %>% filter(Method == 'Observed'), aes(x = t, y = qD), col = "coral2", cex = 2) +
      # for colors
      geom_line(aes(x = t, y = qD,color = Method), cex = 1) +
      geom_point(data = Hills_q0_inc_df %>% filter(Method == 'Observed'), aes(x = t, y = qD, color = Method), cex = 2) +
      # xlab("Number of sampling units") +
      xlab("") +
      ylab("Family Diversity") +
      scale_color_viridis_d() +
      my_ggtheme_p(leg_pos = "bottom") +
      scale_y_continuous(breaks = seq(0, 350, 50),
                         limits = c(0,380)) +
      scale_x_continuous(breaks = seq(0, 22, 2)) +
      scale_linetype_manual(values = c("twodash", "solid"))


```
## S(Observ)

```{r}
N <- nrow(community_matrix_pres)
N
diversity_indices <- Hills_q_0_inc_CCZ$AsyEst

write_csv(
      diversity_indices %>% mutate(category = c("Species Richness", "Shannon div.", "Simpson div"),
                                   ,level = "family",n = 11) %>% 
            select(category,everything()),
      "fam_diversity_indices.csv")


View(diversity_indices)
```


# Pannel figure 


```{r}
# Pannel


bind_rows(Hills_q0_inc_df, Hills_q0_panel_df) %>% 
      mutate(Group = ifelse(is.na(Group), "Family", 
                            ifelse(Group == "Benthic", "Benthic Species","Pelagic Species")
                            ),
             Group = factor(Group, levels = c("Family", "Benthic Species", "Pelagic Species"))) %>% 
      ggplot() +
      geom_line(data = . %>% filter(Method != "Observed"),
                aes(x = t, y = qD, color = Method),
                cex = 1) +
      geom_point(data = . %>% filter(Method == "Observed"),
                 aes(x = t, y = qD, color = Method),
                 cex = 2) +
      facet_wrap(~ Group, scales = "free", ncol = 3) +
      scale_color_viridis_d() +
      scale_y_continuous(breaks = seq(0, 500, 50)) +
      scale_x_continuous(breaks = seq(0, 22, 2)) +
      xlab("Number of Sampling Units") +
      ylab("Richness Estimator") +
      my_ggtheme_p(leg_pos = "right") 


ggsave("diversity_plots.png",
       last_plot(),
       width = 9,
       height = 4)

```

