---
title: "Biodiversity analysis"
author: "Juliano Palacios-Abrantes"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, results='hide', message=FALSE, echo = F}

library(MyFunctions)

packages <- c(
      "readxl", # Read dataframe
      "tidyverse", # for all data wrangling and ggplot
      "janitor",
      "iNEXT",
      "googledrive",
      "googlesheets4"

)

my_lib(packages)
rm(packages)
# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)
```

# Load Data

```{r}
# spp_biodiv_df <- my_path("D", name = "spp_biodiversity_data.csv", read = T)

# fam_biodiv_df <- my_path("D", name = "fam_biodiversity_data.csv", read = T)


sheet_url <- "https://docs.google.com/spreadsheets/d/1ZnycZe_74bB-73LVCrBe56qEmBAa076bv4AFojcArzE/edit?gid=1581131140#gid=1581131140"

spp_biodiv_df <- read_sheet(sheet_url,
                   sheet = "Species_Occurence_Matrix",
                   col_types = "c"
                   ) %>%
      janitor::clean_names() %>% 
      dplyr::mutate(across(cmnfi:species_occurence, ~ as.numeric(.x)))


fam_biodiv_df <- read_sheet(sheet_url,
                   sheet = "Family_Occurence",
                   col_types = "c"
                   ) %>%
      janitor::clean_names()



      

```

# Speceis curve Benthic pelagic 

```{r}

# Transform analysis into function to run by benthic/pelagic group
run_biodiv_analysis <- function(df_subset) {
  
      community_matrix_pres <- df_subset %>%
            pivot_longer(cols = cmnfi:united_states_geological_survey, # Note these must include all columns on data
                         names_to = "site", 
                         values_to = "value") %>%
            group_by(site, scientific_name) %>%
            summarise(presence = ifelse(sum(value) > 0, 1, 0), .groups = "drop") %>%
            reshape::cast(site ~ scientific_name, value = "presence")
      
      
      community_matrix_pres[is.na(community_matrix_pres)] <- 0
      
      inc_freq <- as.incfreq(as.matrix(t(community_matrix_pres)))
      iNEXT(inc_freq, q = 0, datatype = "incidence_freq", nboot = 1, endpoint = 100) # q and qdp is what plots
}

# Split and run analysis by benthic/pelagic group
results_list <- spp_biodiv_df %>%
  group_by(benthic_pelagic) %>%
  group_split() %>%
  set_names(unique(spp_biodiv_df$benthic_pelagic)) %>%
  map(run_biodiv_analysis)


# Extract size-based estimates and tag with group
benthic_df <- as.data.frame(results_list$Benthic$iNextEst$size_based) %>%
  mutate(Group = "Benthic")

pelagic_df <- as.data.frame(results_list$Pelagic$iNextEst$size_based) %>%
  mutate(Group = "Pelagic")

# Combine both into one data frame
Hills_q0_panel_df <- bind_rows(benthic_df, pelagic_df)

  ggplot(Hills_q0_panel_df %>% filter(Method != "Observed")) +
    geom_line(aes(x = t,
                  y = qD, 
                  color = Method),
                  # lty = Method),
              # col = "coral2", 
              cex = 1) +
    geom_point(data = Hills_q0_panel_df %>% filter(Method == "Observed"),
               aes(x = t,
                   y = qD,
                   color = Method),
               # col = "coral2",
               cex = 2) +
    facet_wrap(~ Group, scales = "free_y") +
    xlab("") +
    ylab("Species Diversity") +
    my_ggtheme_p(leg_pos = "right") +
    scale_y_continuous(breaks = seq(0, 3000, 100)) +
    scale_linetype_manual(values = c("twodash", "solid"))

```


## N

In the context of the iNEXT package and biodiversity studies, N typically refers to the total number of sampling units or the total number of individuals observed in the dataset, depending on the type of data being analyzed.

```{r}
N <- nrow(community_matrix_pres)
N
```


## S(Observ)

In the context of biodiversity studies and using the iNEXT package, S(obs) refers to the observed species richness. This is the actual number of species observed in the samples collected. It is an important measure because it represents the baseline diversity observed in your data before any rarefaction or extrapolation.

```{r}
S_obs <- Hills_q_0_inc_CCZ$AsyEst$Observed[1] #544
S_obs

S_se <- Hills_q_0_inc_CCZ$AsyEst$Est_s.e.[1] #544

N <- nrow(community_matrix_pres)
N
diversity_indices <- Hills_q_0_inc_CCZ$AsyEst

write_csv(diversity_indices %>% mutate(category = c("Species Richness", "Shannon div.", "Simpson div"),
                                             ,level = "secies",n = 11) %>% 
                select(category,everything()),
          "spp_diversity_indices.csv")

```

## Extrapolation max

In the context of the iNEXT package, "extrapolation max" refers to the maximum sample size or maximum level of sample coverage to which the rarefaction and extrapolation curves are extended. This parameter is important as it defines how far beyond the observed data the diversity estimates are extrapolated.


# Family curve

```{r}

community_matrix_pres <- fam_biodiv_df %>% 
      gather("site","value",cmnfi:united_states_geological_survey) %>%
      dplyr::group_by(site, family) %>%
      dplyr::summarise(presence = ifelse(sum(value) > 0, 1, 0)) %>%
      reshape::cast(.,  site ~ family, value = "presence")

community_matrix_pres[is.na(community_matrix_pres)] <-  0

inc_freq <- as.incfreq(as.matrix(t(community_matrix_pres)))
Hills_q_0_inc_CCZ <- iNEXT(inc_freq, q = 0, datatype = "incidence_freq", nboot = 1, endpoint = 250) # q and qdp is what plots
Hills_q0_inc_df <- as.data.frame(Hills_q_0_inc_CCZ$iNextEst$size_based)


bio_fam <-
ggplot(Hills_q0_inc_df %>% filter(Method != 'Observed')) +
      geom_ribbon(aes(x = t, ymin = qD.LCL, ymax = qD.UCL), fill = "coral2", alpha = 0.2, show.legend = FALSE) +      
      geom_line(aes(x = t, y = qD, lty = Method), col = "coral2", cex = 1) +
      geom_point(data = Hills_q0_inc_df %>% filter(Method == 'Observed'), aes(x = t, y = qD), col = "coral2", cex = 2) +
      # theme(legend.justification = c(0, 1), 
      #       legend.position = c(.5, .4),
      #       legend.box.margin=margin(c(20, 20, 20, 20)),
      #       legend.margin = margin(1, 1, 1, 1),
      #       legend.spacing.x = unit(0, "mm"),
      #       legend.spacing.y = unit(0, "mm"),
      #       legend.title = element_blank(),
      #       legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) +
      # theme(text=element_text(size=20), #change font size of all text
      #       axis.text=element_text(size=20), #change font size of axis text
      #       axis.title=element_text(size=20), #change font size of axis titles
      #       plot.title=element_text(size=20), #change font size of plot title
      #       legend.text=element_text(size=20), #change font size of legend text
      #       legend.title=element_text(size=20)) + 
      xlab("Number of sampling units") +
      ylab("Family Diversity") +
      my_ggtheme_p(leg_pos = "bottom") +
      # scale_y_continuous(breaks = seq(0, 6000, 1000)) +
      scale_linetype_manual(values = c("twodash", "solid"))


```

```{r}
N <- nrow(community_matrix_pres)
N
diversity_indices <- Hills_q_0_inc_CCZ$AsyEst

write_csv(diversity_indices %>% mutate(category = c("Species Richness", "Shannon div.", "Simpson div"),
                                             ,level = "family",n = 11) %>% 
                select(category,everything()),
          "fam_diversity_indices.csv")

```

```{r}

pp <- gridExtra::grid.arrange(bio_spp,
                        bio_fam,
                        ncol = 1)
ggsave(
      "diversity_plots.png",
      pp,
      width = 7,
      height = 7
)


```

